Supplementary material

### 2022-03-22 submission/manuscript.rmd
fix shiny and rebuild manuscript + supplement

---
title: |
 Individualized treatment effect was predicted best by modeling 
 baseline risk in interaction with treatment assignment
abstract: |
 **Objective**: To compare different risk-based methods for optimal prediction
 of individualized treatment effects. **Study Design and Setting**: We simulated
 RCT data using diverse assumptions for the average treatment effect, a baseline
 prognostic index of risk (PI), the shape of its interaction with treatment
 (none, linear, quadratic or non-monotonic), and the magnitude of
 treatment-related harms (none or constant independent of the PI). We predicted
 absolute benefit using: models with a constant relative treatment effect;
 stratification in quarters of the PI; models including a linear interaction of
 treatment with the PI; models including an interaction of treatment with a
 restricted cubic spline (RCS) transformation of the PI; an adaptive approach
 using Akaike’s Information Criterion. We evaluated predictive performance using
 root mean squared error and measures of discrimination and calibration for
 benefit. **Results**: The linear-interaction model and the RCS-interaction
 displayed robust performance across many simulation scenarios. The RCS-model
 was optimal when quadratic or non-monotonic deviations from a constant
 treatment effect were stronger, and when sample size was larger. The adaptive
 approach required larger sample sizes. Illustrations in the GUSTO-I trial
 confirmed these findings. **Conclusion**: An interaction between baseline risk
 and treatment assignment should be considered to improve treatment effect
 predictions.
author:
  - name: Alexandros Rekkas
    affiliation: 1
  - name: Peter R. Rijnbeek
    affiliation: 1
  - name: David M. Kent
    affiliation: 2
  - name: Ewout W. Steyerberg
    affiliation: 3
  - name: David van Klaveren
    affiliation: 4
address:
  - code: 1
    address: Department of Medical Informatics, Erasmus Medical Center, Rotterdam, The Netherlands
  - code: 2
    address: Predictive Analytics and Comparative Effectiveness Center, Institute for Clinical Research and Health Policy Studies, Tufts Medical Center, Boston, Massachusetts, USA
  - code: 3
    address: Department of Biomedical Data Sciences, Leiden University Medical Center, Leiden, The Netherlands
  - code: 4
    address: Department of Public Health, Erasmus Medical Center, Rotterdam, The Netherlands
numbersections: true
linenumbers: false
journal: arXiv.org
keywords:
  - treatment effect heterogeneity
  - absolute benefit
  - prediction models
output: 
  rticles::elsevier_article:
    keep_tex: true
  bookdown::word_document2:
    reference_docx: reference.docx
geometry: margin=1.0in
date: false
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \usepackage{amsmath}
  - \doublespacing
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \date{}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

@@

>>>>

---
title: |
 Individualized treatment effect was predicted best by modeling 
 baseline risk in interaction with treatment assignment
abstract: |
 **Objective**: To compare different risk-based methods for optimal prediction
 of individualized treatment effects. **Study Design and Setting**: We simulated
 RCT data using diverse assumptions for the average treatment effect, a baseline
 prognostic index of risk (PI), the shape of its interaction with treatment
 (none, linear, quadratic or non-monotonic), and the magnitude of
 treatment-related harms (none or constant independent of the PI). We predicted
 absolute benefit using: models with a constant relative treatment effect;
 stratification in quarters of the PI; models including a linear interaction of
 treatment with the PI; models including an interaction of treatment with a
 restricted cubic spline (RCS) transformation of the PI; an adaptive approach
 using Akaike’s Information Criterion. We evaluated predictive performance using
 root mean squared error and measures of discrimination and calibration for
 benefit. **Results**: The linear-interaction model and the RCS-interaction
 displayed robust performance across many simulation scenarios. The RCS-model
 was optimal when quadratic or non-monotonic deviations from a constant
 treatment effect were stronger, and when sample size was larger. The adaptive
 approach required larger sample sizes. Illustrations in the GUSTO-I trial
 confirmed these findings. **Conclusion**: An interaction between baseline risk
 and treatment assignment should be considered to improve treatment effect
 predictions.
author:
  - name: Alexandros Rekkas
    affiliation: 1
    correspondingauthor: true
  - name: Peter R. Rijnbeek
    affiliation: 1
  - name: David M. Kent
    affiliation: 2
  - name: Ewout W. Steyerberg
    affiliation: 3
  - name: David van Klaveren
    affiliation: 4
address:
  - code: 1
    address: Department of Medical Informatics, Erasmus Medical Center, Rotterdam, The Netherlands
  - code: 2
    address: Predictive Analytics and Comparative Effectiveness Center, Institute for Clinical Research and Health Policy Studies, Tufts Medical Center, Boston, Massachusetts, USA
  - code: 3
    address: Department of Biomedical Data Sciences, Leiden University Medical Center, Leiden, The Netherlands
  - code: 4
    address: Department of Public Health, Erasmus Medical Center, Rotterdam, The Netherlands
numbersections: true
linenumbers: false
journal: arXiv.org
keywords:
  - treatment effect heterogeneity
  - absolute benefit
  - prediction models
output: 
  rticles::elsevier_article:
    keep_tex: true
  bookdown::word_document2:
    reference_docx: reference.docx
geometry: margin=1.0in
date: false
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \usepackage{amsmath}
  - \doublespacing
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \date{}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

@@

# What is new? {-}
* Models including a linear interaction of the prognostic index with treatment
  were a good option for predicting absolute benefit with smaller sample sizes or
  moderately performing prediction models.
* Models including an interaction of treatment with a restricted cubic spline
  transformation using 3 knots performed better in the presence of true
  non-monotonic deviations from a constant relative treatment effect or with
  treatment-related harms.
* Using an AIC adaptive approach to select among a constant effects model, a
  linear interaction model, and a model using restricted cubic spline
  transformations was better suited for larger sample sizes.

# What this adds to what was known? {-}
* Using a constant relative treatment effect for making risk-based
  individualized benefit predictions is not always the best approach.
* Models including linear interaction of treatment with the prognostic index
  have robust performance across a wide variety of settings.
* Increasing the flexibility of the transformation of the prognostic index does
  not improve performance.

# What is the implication and what should change now? {-}
* Assuming a constant relative treatment effect for all patients may lead to
  treatment mistargeting.
* We recommend using a linear interaction of treatment with the prognostic index
  for making individualized risk-based benefit predictions.
* With larger sample size, using an adaptive approach based on AIC to explore
  non-linear interactions of the prognostic index with treatment is a viable
  option.

### 2022-03-22 submission/manuscript.tex
fix shiny and rebuild manuscript + supplement

>>>>

\hypertarget{what-is-new}{%
\section*{What is new?}\label{what-is-new}}
\addcontentsline{toc}{section}{What is new?}

\begin{itemize}
\tightlist
\item
  Models including a linear interaction of the prognostic index with
  treatment were a good option for predicting absolute benefit with
  smaller sample sizes or moderately performing prediction models.
\item
  Models including an interaction of treatment with a restricted cubic
  spline transformation using 3 knots performed better in the presence
  of true non-monotonic deviations from a constant relative treatment
  effect or with treatment-related harms.
\item
  Using an AIC adaptive approach to select among a constant effects
  model, a linear interaction model, and a model using restricted cubic
  spline transformations was better suited for larger sample sizes.
\end{itemize}

\hypertarget{what-this-adds-to-what-was-known}{%
\section*{What this adds to what was
known?}\label{what-this-adds-to-what-was-known}}
\addcontentsline{toc}{section}{What this adds to what was known?}

\begin{itemize}
\tightlist
\item
  Using a constant relative treatment effect for making risk-based
  individualized benefit predictions is not always the best approach.
\item
  Models including linear interaction of treatment with the prognostic
  index have robust performance across a wide variety of settings.
\item
  Increasing the flexibility of the transformation of the prognostic
  index does not improve performance.
\end{itemize}

\hypertarget{what-is-the-implication-and-what-should-change-now}{%
\section*{What is the implication and what should change
now?}\label{what-is-the-implication-and-what-should-change-now}}
\addcontentsline{toc}{section}{What is the implication and what should
change now?}

\begin{itemize}
\tightlist
\item
  Assuming a constant relative treatment effect for all patients may
  lead to treatment mistargeting.
\item
  We recommend using a linear interaction of treatment with the
  prognostic index for making individualized risk-based benefit
  predictions.
\item
  With larger sample size, using an adaptive approach based on AIC to
  explore non-linear interactions of the prognostic index with treatment
  is a viable option.
\end{itemize}

### 2022-03-22 submission/supplement.rmd
fix shiny and rebuild manuscript + supplement

```{r tab2, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
asd <- read_csv(
  here("data/processed/analysisIdsInteractions.csv")
) %>%
  select(-"type")
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable = TRUE,
  align = "r",
  booktabs = TRUE,
  caption = "Scenario settings of the treatment-covariate interaction scenarios.",
  col.names = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    "g0", "g1", "g2", "g5", "g6",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 5,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 6)
```

>>>>

```{r tab2, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
asd <- read_csv(
  here("data/processed/analysisIdsInteractions.csv")
)
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable = TRUE,
  align = "r",
  booktabs = TRUE,
  caption = "Scenario settings of the treatment-covariate interaction scenarios.",
  col.names = c(
    "Scenario",
    "Base",
    "Type",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    "g0", "g1", "g2", "g5", "g6",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 5,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 6)
```
